Curvature approximation on surfaces -- A Discrete Exterior Calculus Approach

### 2014-12-03 article/main.tex
...

  \begin{figure}
    \centering
    \includegraphics[width=.45\textwidth]{bilder/bunnyCurvature.jpg}
    \caption{blub}
    \label{figBunnyCurvature}
  \end{figure}

\section{Discrete Exterior Calculus (DEC)}
  The Discrete Exterior Calculus \citep{hirani, desbrun} defines discrete differential \mh{p}{forms} on a triangulated mesh (simplicial complex).
  For surface meshes, i.e. triangulated orientable \mh{2}{manifolds}, the degree of the discrete \mh{p}{forms} is 0, 1 or 2
  and their are represented by scalars on vertices, edges, triangles or chains of them. 
  Operators for the differential forms, like the exterior derivation \( \exd \) or the hodge star \( \star \), can be approximated by expressions on the discrete
  geometrical structure. 
  E.g. the integral over a triangle of the exterior derivation \( \exd \) for a \mh{1}{form} can be expressed as the integral of the
  \mh{1}{form} over the boundaries edges of the triangle. 
  This follows directly from the Stokes Theorem \citep[Ch. 7]{marsden}.

@@

    \subsubsection{The star operator}
      For a general definition of the star operator \( \star \) see \cite{hirani,desbrun} or \cite[Ch. 7]{siggraphKap7}.
      In our two dimensional well-centered primal mesh case it is enough to define the star operator 
      \( \star: C_{p}(K) \rightarrow C_{2-p}(\csd K) \)
      by 
      \begin{align}
        \label{eqVoronoiCell}
       \star\sigma^{0} &:= \sum_{\sigma^{0} \prec \sigma^{1} \prec \sigma^{2}}
                                                   s_{\sigma^{1} \sigma^{2}} \left[ \sigma^{0}, c(\sigma^{1}), c(\sigma^{2}) \right] \\
       \label{eqVoronoiEdge}
       \star\sigma^{1} &:= \sum_{\sigma^{1} \prec \sigma^{2}}
                                                   s_{\sigma^{2}} \left[ c(\sigma^{1}), c(\sigma^{2}) \right] \\
       \label{eqVoronoiVertex}
       \star\sigma^{2} &:= c(\sigma^{2}) 
      \end{align}
      The factors \( s_{\bullet} \) are signs to prevent orientation properties, like the orientability of the dual mesh.
      \eqref{eqVoronoiCell} describe the Voronoi cell of the primal vertex \( \sigma^{0} \),
      \eqref{eqVoronoiEdge} describe the Voronoi edge of the primal edge \( \sigma^{1} \) (c.p. figure \ref{figExSubdivision})
      and \eqref{eqVoronoiVertex} describe the Voronoi vertex of a primal triangle \( \sigma^{2} \). 
      With the notation \(C_{2-p}( \star K)  \) for the image of \( \star \), 
      the star operator act on \(C_{2-p}( \star K)  \) and map to \( C_{p}(K) \) is for \( p=0,2 \) simply the inverse.
      For \( p=1 \) we must swap the orientation, i.e. \( \star\star\sigma^{1}=-\sigma^{1} \).

@@

    \subsubsection{The discrete Laplace-Beltrami operator}
      The Laplace-Beltrami operator \( \Delta_{B}= * \exd * \exd \) is the special case of the negative Laplace-de Rham operator
      \( -\Delta_{dR} = *\exd *\exd + \exd *\exd * \) for \mh{0}{forms} on a \mh{2}{mainfold} or in general on a even-dimensional manifold.
      To discretize \( \Delta_{B} \) all of work is done above,
      because we can use the definitions \eqref{eqDiscreteHodge} and \eqref{eqDisExD} step by step in the same manner like the smooth exterior
      calculus.
      I.e. for a \mh{0}{form} \( f \) on a vertex \( v \) it results in
      \begin{align}
        \label{eqDisLaplace}
        \left\langle \Delta_{B} f, v \right\rangle
                &= \frac{1}{\left| \star v \right|} \sum_{\sigma^{1}=\left[ v,v_{i} \right]\in K} \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                       \left( f(v_{i}) - f(v) \right) \formpunkt
      \end{align}
      For greater details see \cite{hirani} or \cite{desbrun}. 
      \eqref{eqDisLaplace}, equivalent to the cotan-formular described in \cite{meyer}, is a approximation of order 2 (see \cite{xu})
      and was already earlier published in \cite{arakawa} as a non-DEC consequence.

@@

        A more ``FEM-like'' manner is to decompose \eqref{eqDisLaplace} in parts, which can be computed on all (triangle) elements without
        knowledge of their neighbours.
        For this we formulate the operator on the Voronoi cells to cancel the scale factor \( \left| \star v \right|^{-1} \) 
        in \eqref{eqDisLaplace}, split the sum and use a local vertex indexing on all elements:
        \begin{align} \label{eqVoronoiLaplace}
                \left\langle *\Delta_{B} f , \star v_{i} \right\rangle
                               &= \sum_{\sigma^{1}=\left[ v_{i}, v_{j} \right]} 
                     \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                      \left( f(v_{j}) - f(v_{i}) \right)\\
                               &= \sum_{\substack{\sigma^{2}=\left[ v^{\sigma^{2}}_{0},v^{\sigma^{2}}_{1},v^{\sigma^{2}}_{2} \right]
                              \\
                                               v^{\sigma^{2}}_{0} = v_{i}}}
                  \sum_{l=1,2} C^{\sigma^{2}}_{0,l} 
                          \left( f^{\sigma^{2}}_{l} -  f^{\sigma^{2}}_{0}\right) \formpunkt \label{eqLokalLaplace}
        \end{align}
        \( f^{\sigma^{2}}_{l}\in\R \) are the evaluations of \( f \) on the local vertices \( v^{\sigma^{2}}_{l} \) on the element \(
        \sigma^{2}\).
        Figure \ref{figLocalIndexing} clarify this setting.
        The coefficients \( C^{\sigma^{2}}_{0,l}\in\R \) are generally defined by
        \begin{align}
               C^{\sigma^{2}}_{k,l} &:= C^{\sigma^{2}}_{l,k}
                            := \frac{\left| \star \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] 
                                            \cap \sigma^{2}\right|}
                                   {\left| \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] \right|}\formpunkt
        \end{align}
        \begin{figure}
          \centering
          \includegraphics[width=.45\textwidth]{bilder/tikz/elementSummeKante.eps}
          \caption{Two adjacent (triangle) elements share a common edge \( \sigma^{1}=\left[ v_{i}, v_{j} \right] \).
                   The vertices are local indexed on these elements.}
          \label{figLocalIndexing}
        \end{figure}
        Hence we can extract from \eqref{eqLokalLaplace} a local coefficients matrix
        \begin{align}
          A^{\sigma^{2}} 
              &:= {\footnotesize\begin{bmatrix}
                    -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{02}\right) & C^{\sigma^{2}}_{01} & C^{\sigma^{2}}_{02} \\
                    C^{\sigma^{2}}_{01} & -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{12}\right) & C^{\sigma^{2}}_{12} \\
                    C^{\sigma^{2}}_{02} & C^{\sigma^{2}}_{12} & -\left( C^{\sigma^{2}}_{02} + C^{\sigma^{2}}_{12}\right)
                  \end{bmatrix}} 
        \end{align}
        which holds all computation parts for the local vertices.
        All \( A^{\sigma^{2}} \) can be assembly in a familiar way to get a sparse global matrix 
        \( A\in\R^{N_{\sigma^{0}} \times N_{\sigma^{0}}} \), where \( N_{\sigma^{0}} \) are the number of vertices in \( K \).
        At the end, for \( f_{h}\in\R^{N_{\sigma^{0}}} \), the vector of evaluations for \( f \) 
        (resp. vector of DOFs\footnote{Degree Of Freedom}), 
        \( Af_{h} \) is the result of \eqref{eqVoronoiLaplace} on all global vertices.

@@

    \subsubsection{The discrete gradient}
      The gradient of a scalar valued function \( f \) can be obtained by rising the indices of its exterior derivative, 
      in other words \( \nabla f = \left( \exd f \right)^{\sharp} \).
      Unfortunately, the sharp operator \( \sharp \) depends on the (unknown) metric of the manifold. 
      In \cite{hirani} it is suggested to use the following discrete sharp operator for such a exact \mh{1}{form} \( \exd f \) to get a discrete
      gradient:
      \begin{align}
        \label{eqGradientHeine}
        \left\langle \left( \exd f \right)^{\sharp} , \star\sigma^{2}\right\rangle
          &:= \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(w) \right) \nabla\Phi_{\sigma^{0}}^{\sigma^{2}} \formkomma
      \end{align}
      with a arbitrary vertex \( w\prec\sigma^{2} \) and 
      \( \Phi_{\sigma^{0}}^{\sigma^{2}} \) is the linear test function (hat function) for the vertex \( \sigma^{0} \) 
      restricted to \( \sigma^{2} \). 
      \eqref{eqGradientHeine} gives us a tangential vector of the polytop \( |K| \) on the circumcenter of the triangle \( \sigma^{2} \).
      But in our setting we want to use a gradient on a vertex of the primal mesh, resp. interpreted on the associated Voronoi cell.
      A simple way to get such a discrete gradient, is to average \eqref{eqGradientHeine} over all dual vertices of the Voronoi cell
      related to the primal vertex \( v \).
      \begin{align} \label{eqGradientAverage}
        \left\langle *\overline\nabla f, \star v \right\rangle
            = \sum_{\sigma^{2}\succ v} \left| \star v \cap \sigma^{2} \right|
                 \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(v) \right) \nabla\Phi_{\sigma^{0}}^{\sigma^{2}}
      \end{align}
      The drawback of the discrete average gradient, which we will use in section \ref{secWeingarten}, is that \( |K| \) has no tangential
      space on its vertices, because the polytop is not continuous differentiable in \( v \).
      This makes the situation for a numerical analysis slightly difficult.
      But comparing calculations for several manifolds and functions indicate the consistence of \eqref{eqGradientAverage} to the smooth
      gradient, 
      if the triangulation is fine enough.

@@

\section{Results}

>>>>

  \begin{figure}
    \centering\includegraphics[width=.45\textwidth]{bilder/bunnyCurvature.jpg}
    \centering\includegraphics[width=.45\textwidth]{bilder/horseCurvature.jpg}
    \centering\includegraphics[width=.45\textwidth]{bilder/colourBar.jpg}
    \caption{Surface curvature approximations of the Stanford Bunny and a horse.
             Left: Gaussian curvature \( K \) (square root scaled colours).
             Right: Mean curvature \( H \) (linear scaled colours)}
    \label{figBunnyHorseCurvature}
  \end{figure}

\section{Discrete Exterior Calculus (DEC)}
  The Discrete Exterior Calculus \citep{hirani, desbrun} defines discrete differential \mh{p}{forms} on a triangulated mesh (simplicial complex).
  For surface meshes, i.e. triangulated orientable \mh{2}{manifolds}, the degree of the discrete \mh{p}{forms} is 0, 1 or 2
  and their are represented by scalars on vertices, edges, triangles or chains of them. 
  Operators for the differential forms, like the exterior derivation \( \exd \) or the hodge star \( \star \), can be approximated by expressions on the discrete
  geometrical structure. 
  E.g. the integral over a triangle of the exterior derivation \( \exd \) for a \mh{1}{form} can be expressed as the integral of this
  \mh{1}{form} over the boundaries edges of the triangle. 
  This follows directly from the Stokes Theorem \citep[Ch. 7]{marsden}.

@@

    \subsubsection{The star operator}
      For a general definition of the star operator \( \star \) see \cite{hirani,desbrun} or \cite[Ch. 7]{siggraphKap7}.
      In our two dimensional well-centered primal mesh case it is enough to define the star operator 
      \( \star: C_{p}(K) \rightarrow C_{2-p}(\csd K) \)
      by 
      \begin{align}
        \label{eqVoronoiCell}
       \star\sigma^{0} &:= \sum_{\sigma^{0} \prec \sigma^{1} \prec \sigma^{2}}
                                                   s_{\sigma^{1} \sigma^{2}} \left[ \sigma^{0}, c(\sigma^{1}), c(\sigma^{2}) \right] \\
       \label{eqVoronoiEdge}
       \star\sigma^{1} &:= \sum_{\sigma^{1} \prec \sigma^{2}}
                                                   s_{\sigma^{2}} \left[ c(\sigma^{1}), c(\sigma^{2}) \right] \\
       \label{eqVoronoiVertex}
       \star\sigma^{2} &:= c(\sigma^{2}) 
      \end{align}
      The factors \( s_{\bullet} \) are signs to prevent orientation properties, like the orientability of the dual mesh.
      \eqref{eqVoronoiCell} describe the Voronoi cell of the primal vertex \( \sigma^{0} \),
      \eqref{eqVoronoiEdge} describe the Voronoi edge of the primal edge \( \sigma^{1} \) (c.p. figure \ref{figExSubdivision})
      and \eqref{eqVoronoiVertex} describe the Voronoi vertex of a primal triangle \( \sigma^{2} \). 
      With the notation \(C_{2-p}( \star K)  \) for the image of \( \star \), 
      the star operator, act on \(C_{2-p}( \star K)  \) and map to \( C_{p}(K) \), is for \( p=0,2 \) simply the inverse.
      For \( p=1 \) we must swap the orientation, i.e. \( \star\star\sigma^{1}=-\sigma^{1} \).

@@

    \subsubsection{The discrete Laplace-Beltrami operator} \label{secDisLaplaceBeltrami}
      The Laplace-Beltrami operator \( \Delta_{B}= * \exd * \exd \) is the special case of the negative Laplace-de Rham operator
      \( -\Delta_{dR} = *\exd *\exd + \exd *\exd * \) for \mh{0}{forms} on a \mh{2}{mainfold} or in general on a even-dimensional manifold.
      To discretize \( \Delta_{B} \) all of work is done above,
      because we can use the definitions \eqref{eqDiscreteHodge} and \eqref{eqDisExD} step by step in the same manner like the smooth exterior
      calculus.
      I.e. for a \mh{0}{form} \( f \) on a vertex \( v \) it results in
      \begin{align}
        \label{eqDisLaplace}
        \left\langle \Delta_{B} f, v \right\rangle
                &= \frac{1}{\left| \star v \right|} \sum_{\sigma^{1}=\left[ v,v_{i} \right]\in K} \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                       \left( f(v_{i}) - f(v) \right) \formpunkt
      \end{align}
      For greater details see \cite{hirani} or \cite{desbrun}. 
      \eqref{eqDisLaplace}, equivalent to the cotan-formular described in \cite{meyer}, is a approximation of order 2 (see \cite{xu})
      and was already earlier published in \cite{arakawa} as a non-DEC consequence.

@@

        A more ``FEM-like'' manner is to decompose \eqref{eqDisLaplace} in parts, which can be computed on all (triangle) elements without
        knowledge of their neighbours.
        For this we formulate the operator on the Voronoi cells to cancel the scale factor \( \left| \star v \right|^{-1} \) 
        in \eqref{eqDisLaplace}, split the sum and use a local vertex indexing on all elements:
        \begin{align} \label{eqVoronoiLaplace}
                \left\langle *\Delta_{B} f , \star v_{i} \right\rangle
                               &= \sum_{\sigma^{1}=\left[ v_{i}, v_{j} \right]} 
                     \frac{\left| \star\sigma^{1} \right|}{\left| \sigma^{1} \right|}
                      \left( f(v_{j}) - f(v_{i}) \right)\\
                               &= \sum_{\substack{\sigma^{2}=\left[ v^{\sigma^{2}}_{0},v^{\sigma^{2}}_{1},v^{\sigma^{2}}_{2} \right]
                              \\
                                               v^{\sigma^{2}}_{0} = v_{i}}}
                  \sum_{l=1,2} C^{\sigma^{2}}_{0,l} 
                          \left( f^{\sigma^{2}}_{l} -  f^{\sigma^{2}}_{0}\right) \formpunkt \label{eqLokalLaplace}
        \end{align}
        \( f^{\sigma^{2}}_{l}\in\R \) are the evaluations of \( f \) on the local vertices \( v^{\sigma^{2}}_{l} \) of the element \(
        \sigma^{2}\).
        Figure \ref{figLocalIndexing} clarify this setting.
        The coefficients \( C^{\sigma^{2}}_{0,l}\in\R \) are generally defined by
        \begin{align}
               C^{\sigma^{2}}_{k,l} &:= C^{\sigma^{2}}_{l,k}
                            := \frac{\left| \star \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] 
                                            \cap \sigma^{2}\right|}
                                   {\left| \left[ v^{\sigma^{2}}_{k}, v^{\sigma^{2}}_{l} \right] \right|}\formpunkt
        \end{align}
        \begin{figure}
          \centering
          \includegraphics[width=.45\textwidth]{bilder/tikz/elementSummeKante.eps}
          \caption{Two adjacent (triangle) elements share a common edge \( \sigma^{1}=\left[ v_{i}, v_{j} \right] \).
                   The vertices are local indexed on these elements.}
          \label{figLocalIndexing}
        \end{figure}
        Hence we can extract from \eqref{eqLokalLaplace} a local coefficients matrix
        \begin{align}
          A^{\sigma^{2}} 
              &:= {\footnotesize\begin{bmatrix}
                    -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{02}\right) & C^{\sigma^{2}}_{01} & C^{\sigma^{2}}_{02} \\
                    C^{\sigma^{2}}_{01} & -\left( C^{\sigma^{2}}_{01} + C^{\sigma^{2}}_{12}\right) & C^{\sigma^{2}}_{12} \\
                    C^{\sigma^{2}}_{02} & C^{\sigma^{2}}_{12} & -\left( C^{\sigma^{2}}_{02} + C^{\sigma^{2}}_{12}\right)
                  \end{bmatrix}} 
        \end{align}
        which holds all computation parts for the local vertices.
        All \( A^{\sigma^{2}} \) can be assembly in a familiar way to get a sparse global matrix 
        \( A\in\R^{N_{\sigma^{0}} \times N_{\sigma^{0}}} \), where \( N_{\sigma^{0}} \) are the number of vertices in \( K \).
        At the end, for \( f_{h}\in\R^{N_{\sigma^{0}}} \), the vector of evaluations for \( f \) 
        (resp. vector of DOFs\footnote{Degree Of Freedom}), 
        \( Af_{h} \) is the result of \eqref{eqVoronoiLaplace} on all global vertices.

@@

    \subsubsection{The discrete gradient}
      The gradient of a scalar valued function \( f \) can be obtained by rising the indices of its exterior derivative, 
      in other words \( \nabla f = \left( \exd f \right)^{\sharp} \).
      Unfortunately, the sharp operator \( \sharp \) depends on the (unknown) metric of the manifold. 
      In \cite{hirani} it is suggested to use the following discrete sharp operator for such a exact \mh{1}{form} \( \exd f \) to get a discrete
      gradient:
      \begin{align}
        \label{eqGradientHeine}
        \left\langle \left( \exd f \right)^{\sharp} , \star\sigma^{2}\right\rangle
          &:= \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(w) \right) \nabla_{\R^{3}}\Phi_{\sigma^{0}}^{\sigma^{2}} \formkomma
      \end{align}
      with a arbitrary vertex \( w\prec\sigma^{2} \),
      \( \Phi_{\sigma^{0}}^{\sigma^{2}} \) is the linear test function (hat function) for the vertex \( \sigma^{0} \) 
      restricted to \( \sigma^{2} \) and \( \nabla_{\R^{3}} \) is the usual \( \R^{3} \) gradient in the ambient space. 
      The formulation for the \( \R^{3} \) vector \( \left( \exd f \right)^{\sharp} \) is to read component wise 
      as a \( \R^{3} \)-vectorized scalar formulation.
      \eqref{eqGradientHeine} gives us a tangential vector of the polytop \( |K| \) on the circumcenter of the triangle \( \sigma^{2} \).
      But in our setting we want to use a gradient on a vertex of the primal mesh, resp. interpreted on the associated Voronoi cell.
      A simple way to get such a discrete gradient, is to average \eqref{eqGradientHeine} over all dual vertices of the Voronoi cell
      related to the primal vertex \( v \):
      \begin{align} \label{eqGradientAverage}
        \left\langle *\overline\nabla f, \star v \right\rangle
            = \sum_{\sigma^{2}\succ v} \left| \star v \cap \sigma^{2} \right|
                 \sum_{\sigma^{0}\prec\sigma^{2}} \left( f(\sigma^{0}) - f(v) \right) \nabla_{\R^{3}}\Phi_{\sigma^{0}}^{\sigma^{2}}
      \end{align}
      The drawback of the discrete average gradient, which we will use in section \ref{secWeingarten}, is that \( |K| \) has no tangential
      space on its vertices, because the polytop is not continuous differentiable in \( v \).
      This makes the situation for a numerical analysis slightly difficult.
      But comparing calculations for several manifolds and functions and 
      the results in section \ref{secResults} indicate the consistence of \eqref{eqGradientAverage} to the smooth
      gradient, 
      if the triangulation is fine enough.

      \paragraph{Note on Implementation}
        Like in section \ref{secDisLaplaceBeltrami}, we can deploy a coefficients matrix for the \mh{q}{th} component of 
        \eqref{eqGradientAverage} for a triangle \( \sigma^{2} \):
        \begin{align}
            \begin{bmatrix}
              -\left( C^{\sigma^{2}}_{01q} + C^{\sigma^{2}}_{02q}\right) & C^{\sigma^{2}}_{01q} & C^{\sigma^{2}}_{02q} \\
              C^{\sigma^{2}}_{10q} & -\left( C^{\sigma^{2}}_{10q} + C^{\sigma^{2}}_{12q}\right) & C^{\sigma^{2}}_{12q} \\
              C^{\sigma^{2}}_{20q} & C^{\sigma^{2}}_{21q} & -\left( C^{\sigma^{2}}_{20q} + C^{\sigma^{2}}_{21q}\right)
            \end{bmatrix}
        \end{align}
        with 
        \begin{align}
          C^{\sigma^{2}}_{klq} &:= \left| \star v_{k}^{\sigma^{2}} \cap \sigma^{2} \right|
                                   \frac{\partial}{\partial x^{q}} \Phi^{\sigma^{2}}_{v_{l}^{\sigma^{2}}} \formkomma
        \end{align}
        for \( k,l,q\in\left\{ 0,1,2 \right\} \) 
        and \( \left( x^{0}, x^{1}, x^{2} \right) \) are the ordinary \( \R^{3} \) coordinates.

@@

\section{Results} \label{secResults}

  \begin{align}
    \label{eqSphere}
    \varphi\left( x,y,z \right) &:= x^{2} + y^{2} + z^{2} - 1 \\
    \label{eqEllipsoid}
    \varphi\left( x,y,z \right) &:= \left( 3x \right)^{2} + \left( 6y \right)^{2} + \left( 2z \right)^{2} - 9\\
    \label{eqQuartic}
    \varphi\left( x,y,z \right) &:= \left( x - z^{2} \right)^{2} + \left( y - z^{2} \right)^{2} + z^{2} - 1
  \end{align}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/sphere/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/sphere/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/sphere/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/sphere/LMaxH.eps}
    \end{minipage}
    \caption{Sphere \eqref{eqSphere}: Relative \( L_{2} \) (top) resp. \( L_{\infty} \) (bottom) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figSphere}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/H.jpg}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/ellipsoid/LMaxH.eps}
    \end{minipage}
    \caption{Ellipsoid \eqref{eqEllipsoid}: Top row: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).
                              Middle and bottom row: Relative \( L_{2} \) resp. \( L_{\infty} \) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figEllipsoid}
  \end{figure}

  \begin{figure}
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/K.jpg}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/H.jpg}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/L2K.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/L2H.eps}
    \end{minipage}\\
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/LMaxK.eps}
    \end{minipage}\hfill
    \begin{minipage}[htp]{.23\textwidth}
      \centering
      \includegraphics[width=0.99\textwidth]{bilder/quartic/LMaxH.eps}
    \end{minipage}
    \caption{Quartic Surface \eqref{eqQuartic}: Top row: Gaussian curvature \( K \) (left) and mean curvature \( H \) (right).
                              Middle and bottom row: Relative \( L_{2} \) resp. \( L_{\infty} \) error for \( K \) (left) and
                                                     \( H \) (right) for different \( h \) in a log-log-plot.}
    \label{figQuartic}
  \end{figure}
